---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Let's simple read file, convert data into data format  and see result.

```{r}
DateType  <- "%Y-%m-%d"
FitbitData <- read.csv(unz("activity.zip", "activity.csv"), header = T)
FitbitData$date <- as.Date(FitbitData$date, DateType)
#FitbitData$date <- as.factor(FitbitData$date)
summary(FitbitData)
```


## What is mean total number of steps taken per day?

For this part I've calculated total numbers of steps for each day.
```{r}
a <- aggregate(FitbitData$steps, by=list(Category=FitbitData$date), FUN=sum)
```
So, mean and median value are:
```{r}
print(paste0("Mean: ", as.character(mean(a$x, na.rm = T))))
print(paste0("Median: ", as.character(median(a$x, na.rm = T))))
```
And plot histogram:

```{r}
#summary(a)
hist(a$x, main = "Total number of steps per day", xlab = "Steps per day", col = "cadetblue2")
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
